A multi-scale story of the diffusion of a new technology: the web

### 2024-06-11 paper/submission1/tranos2023.tex
pre-ggplot

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, sharp corners, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related \emph{future} technologies. As per
\citet{leibowicz2016representing}, historical studies agree on the fact
that ``technologies diffuse at different times, at different rates, and
to different extents in different places, and can be significantly
influenced by policies \citep{victor1993}''. \citet{meade2021modelling}
highlight that a variety of actors have a direct interest in gaining
such knowledge including network equipment suppliers; network operators,
regulatory and local authorities. These processes and their effects very
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places \citep{leibowicz2016representing}. Despite the spatial
heterogeneity of such diffusion mechanisms and the policy relevance,
there are very limited attempts in the literature to analyse the
diffusion of new digital technologies at a detailed geographical level.

@@

These ideas were firstly introduced by Torsten Hägerstrand and his
thesis entitled `Innovation Diffusion as a Spatial Process'
\citep{hagerstrand1968innovation}. Hägerstrand was the first one to
identify diffusion as a geographical process. The starting point was the
idea that diffusion is based on passing information through social
networks, which themselves tend to be defined by geography. Hence, he
identified the `neighbourhood' effect of how information, and
consequently, innovation diffuse. He used agricultural innovations to
test and model his ideas using Monte Carlo simulations. Hägerstrand also
incorporated the role of hierarchy and how some phenomena maybe firstly
adopted in larger cities and then diffuse to second tier ones. This is a
sequential instead of a simultaneous process, which resembles the
`lead-lag' spatial acceleration effect in market research
\citep{bento2018time, PERES201091}. Hägerstrand maybe more widely known
for highlighting the role time plays in the diffusion of innovations: an
early-pioneering period, a middle fast accelerating period and a final
saturation one {[}\citet{morrill2020spatial}.

The temporal dimension was further explored by Everett Rogers and his
seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
Rogers being a sociologist his work focused not on the diffusion of
innovations over space and time, but instead on the adoption of new
technologies and innovations by individuals and the individual
mechanisms that drive the decisions behind adoption. He identified five
groups of individuals regarding their adoption speed: innovators, early
adopters, early majority, late majority and laggards. Key mechanism of
diffusion and adoption is always communication and how knowledge is
transferred within a social system. Therefore, all approaches agree on
the S-shaped diffusion and cumulative adoption pattern
\citep{grubler1990rise}.

@@

The remaining of this section reviews empirical studies which analysed
the diffusion on new technologies over space and time. Although the
spatial dimension is present in most of the following studies, the level
of spatial detail is almost always more coarse than the one adopted
here. To begin with \citet{beardsell1999spatial} studied the evolution
of the computer industry in \(317\) US metro areas during the
\(1977-1992\) period using employment data. Their analysis indicated
that the relative size distribution holds for urban computer employment
and also urban heterogeneity is essential in explaining this
distribution. \citet{bednarz2020pulled} focused on wind turbines and
modelled their spatial diffusion across \(402\) German regions during
\(1970-2015\). Their key finding is that local demand than local supply
was the main driving factor.

At a global scale \citet{perkins2005international} analyzed if the
diffusion rate of new technologies is driven by a latecomer advantage
and the engagement with the global economy via foreign direct
investments and trade. Their results illustrate that indeed latecomers
and developing countries experience diffusion of new technologies more
rapidly than early adopter and developed countries. At the same scale,
\citet{perkins2011internet} explored whether the adoption of previous
communication technologies that is mail, telegrams and telephones was
shaped by similar socioeconomic factors as the internet. Their results
indicated common patterns regarding the drivers behind the adoption of
different communication technologies.

@@

\hypertarget{sec3.1}{%
\subsection{Results}\label{sec3.1}}

@@

\begin{itemize}
\item
  Moran's I: for OA and LAD over time \textbf{TODO} add 0s
\item
  LISA maps: for OA and LAD over time More and less expected clusters.
  Different scales show different results
\item
  Website density regressions: for OA and LAD over time \textbf{TODO}
  add 0s Similar pattern
\end{itemize}

@@

\begin{itemize}
\item
  S for LADs per firm and OA. \textbf{TODO}: OA per firm? fix firms over
  time.
\item
  Fast and slow LAs map. There is clustering
\item
  \textbf{TODo} check s\_uk\_firm\_11
\item
  ranks: there is stability and movement
\end{itemize}

\hypertarget{sec3.2}{%
\subsection{3.2 Case study approach}\label{sec3.2}}

\hypertarget{sec4}{%
\section{Digital technologies and spatial structure: a global
view}\label{sec4}}

\hypertarget{sec5}{%
\section{The impact of ICT on the US and the UK spatial
structures}\label{sec5}}

\hypertarget{sec6}{%
\section{Discussion and conclusions}\label{sec6}}

>>>>

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, frame hidden, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners]}{\end{tcolorbox}}\fi

@@

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related \emph{future} technologies. As per
\citet{leibowicz2016representing}, historical studies agree on the fact
that ``technologies diffuse at different times, at different rates, and
to different extents in different places, and can be significantly
influenced by policies'' \citep{victor1993}. \citet{meade2021modelling}
highlight that a variety of actors have a direct interest in gaining
such knowledge including network equipment suppliers; network operators,
regulatory and local authorities. These processes and their effects vary
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places \citep{leibowicz2016representing}. Despite the spatial
heterogeneity of such diffusion mechanisms and the policy relevance,
there are very limited attempts in the literature to analyse the
diffusion of new digital technologies at a detailed geographical level.

@@

These ideas were firstly introduced by Torsten Hägerstrand and his
thesis entitled `Innovation Diffusion as a Spatial Process'
\citep{hagerstrand1968innovation}. Hägerstrand was the first one to
identify diffusion as a geographical process. The starting point was the
idea that diffusion is based on passing information through social
networks, which themselves tend to be defined by geography. Hence, he
identified the `neighbourhood' effect of how information, and
consequently, innovation diffuse. He used agricultural innovations to
test and model his ideas using Monte Carlo simulations. Hägerstrand also
incorporated the role of hierarchy and how some phenomena maybe firstly
adopted in larger cities and then diffuse to second tier ones. This is a
sequential instead of a simultaneous process, which resembles the
`lead-lag' spatial acceleration effect in market research
\citep{bento2018time, PERES201091}. Hägerstrand is more widely known
though for highlighting the role time plays in the diffusion of
innovations: an early-pioneering period, a middle fast accelerating
period and a final saturation one \citep{morrill2020spatial}.

The temporal dimension was further explored by Everett Rogers and his
seminal work on `Diffusion of Innovations' \citep{rogers2010diffusion}.
Rogers being a sociologist, his work focused not on the diffusion of
innovations over space and time, but instead on the adoption of new
technologies and innovations by individuals and the individual
mechanisms that drive the decisions behind adoption. He identified five
groups of individuals regarding their adoption speed: innovators, early
adopters, early majority, late majority and laggards. Key mechanism of
diffusion and adoption is always communication and how knowledge is
transferred within a social system. Therefore, all approaches agree on
the S-shaped diffusion and cumulative adoption pattern
\citep{grubler1990rise}.

@@

The remaining of this section reviews empirical studies which analysed
the diffusion on new technologies over space and time. Although the
spatial dimension is present in most of the following studies, the level
of spatial detail is almost always more coarse than the one adopted
here. \citet{beardsell1999spatial} studied the evolution of the computer
industry in \(317\) US metro areas during the \(1977-1992\) period using
employment data. Their analysis indicated that the relative size
distribution holds for urban computer employment and also urban
heterogeneity is essential in explaining this distribution. in a recent
study, \citet{bednarz2020pulled} focused on wind turbines and modelled
their spatial diffusion across \(402\) German regions during
\(1970-2015\). Their key finding is that local demand than local supply
was the main driving factor.

At a global scale \citet{perkins2005international} explored whther the
diffusion rate of new technologies is driven by a latecomer advantage
and the engagement with the global economy via foreign direct
investments and trade. Their results illustrate that indeed latecomers
and developing countries experience diffusion of new technologies more
rapidly than early adopter and developed countries. At the same scale,
\citet{perkins2011internet} explored whether the adoption of previous
communication technologies that is mail, telegrams and telephones was
shaped by similar socioeconomic factors as the internet. Their results
indicated common patterns regarding the drivers behind the adoption of
different communication technologies.

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{old-ideas}{%
\subsection{Old ideas}\label{old-ideas}}

@@

\begin{itemize}
\item
  Moran's I: for OA and LAD over time
\item
  LISA maps: for OA and LAD over time More and less expected clusters.
  Different scales show different results
\item
  Website density regressions: for OA and LAD over time \textbf{TODO}
  add 0s Similar pattern
\end{itemize}

@@

\begin{itemize}
\item
  S for LADs per firm and OA.
\item
  Fast and slow LAs map. There is clustering
\item
  s\_uk\_firm\_11
\item
  ranks: there is stability and movement
\end{itemize}

\hypertarget{new-structure}{%
\subsection{New structure}\label{new-structure}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples
\item
  ranks: there is stability and movement
\item
  Neighbourhood effect: diffusion proceeds outwards from innovation
  centers, first ``hitting'' nearby rather than far-away locations
  (Grubler
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\setcounter{enumi}{1989}
\tightlist
\item
\end{enumerate}

\begin{itemize}
\item
  Moran's I: for OA and LAD over time
\item
  LISA maps: for OA and LAD over time More and less expected clusters.
  Different scales show different results
\item
  Website density regressions: for OA and LAD over time
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Hierarchy effect: from main centers to secondary ones -- central
  places
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  RF
\end{enumerate}

\begin{itemize}
\tightlist
\item
  ideal: (i) train RF for all years and all (1) LADs and (2) OAs with
  CAST and report metrics. (ii) train for all years and all but one
  region for (1) LADs and (2) OAs to predict to the holdout region.
  Reports predictions as region similarities.
\end{itemize}

\hypertarget{rf-results}{%
\subsubsection{RF results}\label{rf-results}}

The next section incorporates the above discussed spatial drivers of the
diffusion of web technologies into the same modelling framework. The aim
is to use variables depicting these spatial processes in order to
predict the diffusion of web technologies in the UK over space and time
and across different scales. Specifically, four different models are
estimated. Firstly, all the data points at the OA and LAD are utilised
in order to build RF models and assess their capacity to predict the
adoption of web technologies. These two models will reveal the
predictive capacity of the spatial processes and also allow to see how
the importance of such variables changes across scales. The next two
models will be again trained on web difusion at the two working scales:
OA and LAD. However, instead of using all the data points, the OA and
the LAD from one of the 12 UK regions are hold out and then the trained
model is used to predict web adoption in the OA or the LAD of the
held-out region. This process takes place recursively for all UK
regions. The difference in the predictive capacity of the different
samples will reveal how different are these spatial process across
regions and, importantly, at different scales.

It needs to be highlighted here that the cross-validation for all models
presented here is spatially and temporally sensitive. Instead of using
10 random folds, we employ the \texttt{CAST} package which allows holds
back data points from specific years and spatial units and use them for
testing in order to estimate the model performance
\citep{meyer2018improving}.

The models need to include variables that capture the three processes
that the relevant literature and the descriptive analysis presented in
the previous section highlighted. Namely, the models capture: (i) a
hierarchy effect with diffusion running from main centres to secondary
ones, (ii) a neighborhood effect according to which diffusion first hits
nearby locations, and (iii) the rather canonical pattern of the
diffusion over time as reflected in the S-shaped pattern in the
cumulative level of adoption.

To capture the hierarchy effect the models include as predictors a one
year lag of website density in London -- the largest city in the UK, a
one year lag of the website density in the nearest city and the same for
the nearest retail centre. Due to the small sizes of the retail centres,
the latter is only relevant for the OA-level models. In addition, the
models include the distance to London, the nearest city and the nearest
retail centre. The underlying logic is that the level of website
adoption in a spatial unit depends on the level of the adoption in
places further up in the urban hierarchy the previous year. To depict
the neighbourhood effect, the web density of the neighbouring spatial
units in the previous year is calculated. Again, the underpinning
rationale is that the level of web adoption within a spatial unit
depends on the level of web adoption in the neighbouring spatial units
the year before. This is the `hitting nearby locations first' argument.
Therefore, the spatial and temporal lag of the website density in LAD
and OA is calculated. Lastly, the time effect which is reflected on the
S curve for the cumulative adoption is captured by trend variable.
Hence, all four model will follow the following generic form
\ref{model}:

\begin{align} \label{model}
Website\,Density_{t} \sim Distance\,London +
Website\,density\,London_{t-1} +\notag\\
Distance\,Nearest\,City +
Website\,density\,Nearest\,City_{t-1} +\notag\\
Distance\,Nearest\,Retail_{i} +
Website\,density\,Nearest\,Retail_{t-1} +\notag\\
W*\, Website\,density_{t-1} +\notag\\ 
year_{t}
\end{align}

To assess the predictive capability of the model, three broadly utilised
metrics are employed: the coefficient of determination (R squared), mean
absolute error (MAE) and root mean square error (RMSE):

\begin{align}
R^2 = 1 - \frac{\sum_{k} (y_{k} - \hat{y_{k}})^2} {\sum_{k} (y_{k} - \overline{y_{k}})^2} \label{eq:rsquared}
\end{align}

\begin{align}
MAE = \frac{1}{N} \sum_{k = 1}^{N} |\hat{y_{k}} - y_{k}| \label{eq:mae}
\end{align}

\begin{align}
RMSE =  \sqrt{\frac{\sum_{k = 1}^{N} (\hat{y_{k}} - y_{k})^2} {N}} \label{eq:rmse}
\end{align}

\(y_{k}\) is the \(k^{th}\) observation of the dataset, which consists
of \(N\) observations in total. \(\hat{y_{k}}\) is the \(k_{th}\)
predicted value for the dependent variable and \(\overline{y_{k}}\) is
the average value of \(y\). The last two metrics are expressed in the
same units as the dependent variable -- websites per firm for the LAD
modes and the number of websites for the OA models -- while the first
one is the coefficient of determination between the observed and the
predicted values of website adoption. Regarding \(MAE\), it is the
absolute difference between the observed and the predicted website
adoption. While \(MAE\) does not penalise for large errors, \(RSME\)
does so as it is proportional to the squared difference between the
observed and the predicted trade flows. Hence larger errors weigh more
for \(RMSE\) \citep{pontius2008components}.

Table \ref{table.metrics.all} presents the model performance for the
first set on models, for which all data points are employed for training
and testing via cross validation. The first one is trained and tested on
374 LAD and the second on 232,296 OA for a 16 year period (1997-2012).
The results are remarkably good considering that the are the outcome of
space and time sensitive CV, so the the model does not suffer from
overfit. At the LAD level we the model predicts 81\% of the variation of
website density. Both error metrics indicate that the model error is
\((\frac{1}{20} - \frac{1}{30})\) of a website per firm. At the OA this
score drops down to 21\%. Considering its granularity, this is still a
remarkable performance. To contextualise it, the model results in a MAE
of one website for areas small enough to host less than 140
households.\footnote{According to the Office for National Statistics,
  80\% of OA in England and Wales host 110-139 households,
  \href{https://www.ons.gov.uk/census/2001censusandearlier/dataandproducts/outputgeography/outputareas}{www.ons.gov.uk}.}
Because of the small size of the spatial units, the distribution is
highly skewed and a significant part of them is not linked to any
websites. In 1997 only 1\% of the UK OA were associated with a website.
This should not come as a surprise as this was the very beginning of the
commercial internet and any activities with a digital footprint were
concentrated in a handful of areas. This was clearly illustrated in
Section \textbf{ADD}. At the end of the study period almost half of the
UK OA were not associated with a website. Again, given the granularity
of the data this should not come as a surprise. Despite the large
heterogeneity in website adoption, the model reached an R-Squared of
0.2.

\begin{longtable}[]{@{}llll@{}}
\caption{Model metrics \{\#table.metrics.all\}}\tabularnewline
\toprule\noalign{}
& RMSE & RSquared & MAE \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& RMSE & RSquared & MAE \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Local Authorities & 0.032 & 0.810 & 0.019 \\
Output Areas & 5.000 & 0.205 & 1.047 \\
\end{longtable}

Figures \ref{var.imp.LAD} and \ref{var.imp.OA} plot the importance of
the difrerent predictors. When the focus is on the LAD, the website
density in the nearest city, in London and on the neighbouring LADs the
year before are the most important predictors. They are followed by the
yearly trend, while the spatial configuration as reflected on distances
to London or the nearest city only play a minor role. This can be
attributed to the rather coarse spatial scale of analysis. Nevertheless,
all previously discussed forces are at play in the diffusion of web
technologies at the LAD: the first two predictors depict the
hierarchical effect, the spatial and temporal lag of website density the
neighbourhood effect and the yearly trend the time-sensitive cumulative
adoption pattern.

When the much more granular scale of OA is adopted, the picture is
reversed. The most important predictors are the three distance variables
to London, the nearest city and the nearest retail centre. They stil
depict the hierarchical effect, but proximity to the different
population centres is more important than their lagged web densities in
predicting website diffusion. The neighbouring effect is less important
at this scale. What is interesting is the almost negligible role of the
yearly trend and London's website density. While the former probably
illustrates the large heterogeneity in how web technologies have been
adopted at this very fine scale, the later highlights that the
importance of past web adoption rates in large population centres is
surpassed by proximity to them and soatial configuration at this scale.

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/rf/figures/varimp_LA.png}

}

\caption{\label{var.imp.LAD}Variable importance, LAD}

\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{../../outputs/rf/figures/varimp_OA.png}

}

\caption{\label{var.imp.OA}Variable importance, OA}

\end{figure}

Table \ref{table.regions} presents the results of the recursive hold out
models which aim to highlight potential regional heteorigeinty of the
web diffusion spatial mechanisms. To begin with, as highlighted before,
there is a difference of magnitude of one order between the LAD and the
OA prediction errors, which is aligned with previous results that
employed all data points. What is of interest here is the regional
comparison. Table \ref{table.regions} illustrates some striking
similarities, but also a few significant differences. The regions the
web diffusion of which is better predicted using models trained in the
rest of the country are the same despite the scale of analysis: South
East, Wales, Yorkshire and The Humber and the North East of England. In
other words, these are the regions whose spatial diffusion of web
technologies is closer to the country's average. Despite the consistency
across scales, this is a diverse set of regions: \textbf{ADD
CHARACTERISTICS}.

At the other end of the spectrum, Scotland's and the North West's web
diffusion mechanisms are consistently diverging from the country's
average. This should not come as a surpise as these regions are
characterised of high levels of rurality and remoteness. Similarly,
London diffusion mechanisms diverge from the country's average and this
is consistent across scales. London's uniqueness in UK's urban system
and economy is also reflected in the spatial diffusion mechanisms of web
technologies within its LAD and OA. It needs to be highlighted though
that the difference between the RSquared of LAD and OA is more that an
order of magnitude signaling how difficult is to predict diffusion at
such a small spatial scale. Northern Ireland is an interesting case.
While it ranks at the bottom of the scale when the models are trained
and tested on LAD data, when the modelling adopts the more granular OA
scale, the spatial mechanisms that shape the web diffusion within this
region appear to be closer to the country's average. At this scale,
proximity or lack of in relative terms to the rest of the country become
a less important and the internal to the region spatial structure
predictors start playing a more important role \textbf{CHECK NI OA}.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3731}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1940}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1343}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1791}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1194}}@{}}
\caption{Regional differences\label{table.regions}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Region
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
RSquared LAD
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Rank LAD
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
RSquared OA
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Rank OA
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Region
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
RSquared LAD
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Rank LAD
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
RSquared OA
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Rank OA
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
South East & 0.947 & 1 & 0.134 & 2 \\
Wales & 0.916 & 2 & 0.131 & 3 \\
Yorkshire and The Humber & 0.906 & 3 & 0.144 & 1 \\
North East & 0.895 & 4 & 0.128 & 4 \\
West Midlands & 0.883 & 5 & 0.070 & 9 \\
East Midlands & 0.882 & 6 & 0.088 & 8 \\
East of England & 0.876 & 7 & 0.106 & 6 \\
South West & 0.864 & 8 & 0.117 & 5 \\
London & 0.805 & 9 & 0.055 & 10 \\
Scotland & 0.770 & 10 & 0.035 & 11 \\
North West & 0.664 & 11 & 0.017 & 12 \\
Nortern Ireland & 0.576 & 12 & 0.101 & 7 \\
\end{longtable}

\hypertarget{discussion-and-conclusions}{%
\section{Discussion and conclusions}\label{discussion-and-conclusions}}

### 2024-06-11 presentations/web_diffusion.Rmd
pre-ggplot

---
title: 'A multi-scalar story of the diffusion of a new technology: the web'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: bibliography.bib
biblio-style: "apalike"
css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

@@

-   **Granular** and **multi-scale** spatial perspective

@@

-   Descriptive statistics, ESDA & density regressions

@@

-   Websites in Output Areas (c. 200,000)

@@

|                   | RMSE  | RSquared | MAE   |
|-------------------|-------|----------|-------|
| Local Authorities | 0.032 | 0.810    | 0.019 |
| Output Areas*      | 5.000 | 0.205    | 1.047 |

<small>
*No retail centre variables for the Local Authority models
</small>

@@

## Regional similarities for Local Authorities

```{r similarity_LA, eval = T, message=FALSE}
path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
read.csv(path.table) %>% 
  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
  kable()
```

## Regional similarities for Output Areas

```{r similarity_OA, out.height= "70%", out.width="70%", message=FALSE}
path.table <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")
read.csv(path.table) %>% 
  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
  kable()
```

@@

<small>
*No retail centre variables for the Local Authority models
</small>

@@

```{r growth LA}
path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
read.csv(path.table) %>% 
  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
  kable()
```

@@

-   Established technological diffusion drivers still apply

@@

-   Heterogeneity increases with resolution

-   Things to do: OA?

>>>>

---
title: 'A multi-scale story of the diffusion of a new technology: the web'
# date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
author: |
  |
  | Emmanouil Tranos
  |
  | <small>University of Bristol, Alan Turing Institute</small>
  | <small>e.tranos@bristol.ac.uk, [\@EmmanouilTranos](https://twitter.com/EmmanouilTranos), [etranos.info](https://etranos.info/)</small>
output: 
  revealjs::revealjs_presentation:
    theme: simple
    self_contained: true
    reveal_options:
      width: 1150
      height: 720
bibliography: bibliography.bib
biblio-style: "apalike"
css: style.css
editor_options: 
  markdown: 
    wrap: 72
---

@@

-   Granular and multi-scale **spatial processes**

@@

-   Descriptive statistics & ESDA

@@

-   Websites in Output Areas (c. 230,000)

@@

|                   | RMSE  | RSquared | MAE   |
|-------------------|-------|----------|-------|
| Local Authorities* | 0.032 | 0.810    | 0.019 |
| Output Areas       | 5.000 | 0.205    | 1.047 |

<small>
*No retail centre predictors for the Local Authority models
</small>

@@

## Regional similarities

```{r similarity, eval = T, message=FALSE}
path.lad <- paste0(path, "/outputs/rf/figures/test_regions_LA.csv")
path.oa <- paste0(path, "/outputs/rf/figures/test_regions_OA.csv")

lad <- read.csv(path.lad) %>% 
  rename('RSquared LAD' = Rsquared)

oa <- read.csv(path.oa) %>% rename(Region = test.region,
                                   'RSquared OA' = Rsquared)

lad %>% left_join(oa) %>% 
  mutate('Rank OA' = rank(desc(`RSquared OA`)),
         'Rank LAD' = rank(desc(`RSquared LAD`))) %>% 
  relocate(5, .after = 2) %>% 
  arrange(`Rank LAD`) %>% 
  kable(digits=3)
```

@@

<small>
*No retail centre predictors for the Local Authority models
</small>

@@

```{r growth LA}
path.table <- paste0(path, "/outputs/rf/figures/test_regions_LA_growth.csv")
read.csv(path.table) %>% 
  mutate(Rsquared = round(Rsquared, digits= 3)) %>% 
  arrange(desc(Rsquared)) %>% 
  kable()
```

@@

-   Established technological diffusion spatial processes still apply

@@

-   Regional heterogeneity
